---
title: "Law School Admissions"
format: html
execute: 
  echo: false
---

```{r}
#| message: false
library(tidyverse)
library(rstanarm)
```

```{r}
#| message: false

# First pass, I want a variable, accepted, that is TRUE for everyone who was
# accepted, either regular or off the wait-list. Since there were only 8 such
# students, their status can't effect the results much.

x <- read_csv("BU_2020.csv") |> 
  mutate(accepted = if_else(result %in% c("Accepted", "WL, Accepted"), TRUE, FALSE))
```

As we would expect, accepted students have high LSAT scores and grade point averages.

```{r}
x |> 
  summarize(total = n(),
            avg_lsat = mean(lsat),
            avg_gpa = mean(gpa2),
            .by = accepted)
```

Among all applicants, under-represented minorities, have lower LSAT scores and GPAs.

```{r}
x |> 
  summarize(total = n(),
            avg_lsat = mean(lsat),
            avg_gpa = mean(gpa2),
            .by = urm2)
```

Looking at the cross-tabs, accepted URMs have lower scores/grades than other accepted students.

```{r}
x |> 
  summarize(total = n(),
            avg_lsat = mean(lsat),
            avg_gpa = mean(gpa2),
            .by = c(accepted, urm2))
```

All of this suggests that LSAT, GPA and URM-status plays a role in admissions. A simple Bayesian logistic model confirms that intuition.

```{r}
mod_obj_1 <- stan_glm(accepted ~ lsat + gpa2 + urm2,
                    family = binomial(),
                    data = x,
                    refresh = 0,
                    seed = 9)

mod_obj_1 
```

The three predictors all have the expected sign and are all highly statistically significant. Are the coefficients robust to other modelling choices? Tough to say. I found no evidence that interaction terms are important. However, if we defines `accepted` to include anyone accepted (either regular or off the waitlist) or placed on the waitlist at the beginning, we get a different model.


```{r}
y <- read_csv("BU_2020.csv") |> 
  mutate(accepted = if_else(result %in% c("Accepted", "WL, Accepted", "Waitlisted"), TRUE, FALSE))
  


mod_obj_2 <- stan_glm(accepted ~ lsat + gpa2 + urm2,
                    family = binomial(),
                    data = y,
                    refresh = 0,
                    seed = 9)

mod_obj_2
```

All the coeffients have the correct sign and are significant. But all are *smaller*. For me, this is a sign that the waitlist decision is much noisier, subject to various weirdnesses. Including these applicants as accepted *dilutes* the signal, as if we just added random noice to the process. So, I will stick with the first model.



